When does vapor pressure deficit drive or reduce evapotranspiration?

### 2017-12-13 doc/agu_2017/vpd.tex
pretty much done with slides, just proofread, see if columbia template, maybe change to kPa, and add some more extras

\usepackage[english]{babel}
\usepackage[export]{adjustbox}
 % \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}
\usepackage{tikz}
\usepackage{xcolor}
\graphicspath{{figures/}}
\DeclareMathOperator{\sign}{sign}

@@

\begin{frame}{Consequences of theory - ``sign'' term}
  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ \textcolor{magenta}{\textbf{uWUE }}}} \left( \frac{2 \textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD}}{2 (\textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD})^2}\right) \right)\]
  \begin{itemize}
  \item $c_p$ and $R*$ are constants
  \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
  \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (i.e. grass, crops, deciduous forest, evergreen needleleaf forest, shrub)
  \end{itemize}
  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
\end{frame}

@@

\section{Test Theory}
\begin{frame}{The theory seems nice, but we should test with data!}
    \begin{itemize}
    \item Introduce a free uncertainty parameter $\sigma$ to Penman Monteith:
      \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \boldsymbol{\sigma} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
    \item At each observation from FLUXNET (56 sites across globe) calculate a unique $\sigma$:
      \small
      \[\sigma = - \frac{g_a \gamma c_s \sqrt{VPD} P }{ \left(\text{ ET } ( \Delta + \gamma) - \Delta R - g_a \rho_a c_p VPD\right) 1.6 \; R\; T\; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})}\]
      \normalsize
    \item Then propagate uncertainty forward by including $\sigma$ in the derivative:
    \end{itemize}
      \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \sigma \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]

@@

\begin{frame}{Extra slide - is theory VPD$_{crit}$ optimum?}
  \includegraphics[width=\textwidth]{for_agu.pdf}
\end{frame}

>>>>

\usepackage[english]{babel}
\usepackage[export]{adjustbox}
 % \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{relsize}
\usepackage[absolute,overlay]{textpos}
\graphicspath{{figures/}}
\DeclareMathOperator{\sign}{sign}

@@

\begin{frame}{Consequences of theory - ``sign'' term}
  \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ \textcolor{magenta}{\textbf{uWUE }}}} \left( \frac{2 \textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD}}{2 (\textcolor{magenta}{\boldsymbol{g1}} + \sqrt{VPD})^2}\right) \right)\]
  \begin{itemize}
  \item $c_p$ and $R*$ are constants
  \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
  \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (e.g. grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
  \end{itemize}
  So within each plant type, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is essentially just a function of VPD!
\end{frame}

@@

\begin{frame}{``Sign'' term as a function of VPD and PFT}
  \begin{figure}
      \begin{tikzpicture}
    \node[anchor=south west,inner sep=0] (image) at (0,0) {\adjincludegraphics[width=3.3in, trim={0 {0.51\height} 0 0}, clip]{fig05_pet.pdf}};
        \begin{scope}[x={(image.south east)},y={(image.north west)}]
        %% next four lines will help you to locate the point needed by forming a grid. comment these four lines in the final picture.↓
       % \draw[help lines,xstep=.1,ystep=.1] (0,0) grid (1,1);
       % \draw[help lines,xstep=.05,ystep=.05] (0,0) grid (1,1);
       % \foreach \x in {0,1,...,9} { \node [anchor=north] at (\x/10,0) {0.\x}; }
       % \foreach \y in {0,1,...,9} { \node [anchor=east] at (0,\y/10) {0.\y};}
       %% upto here
        % \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,0.3in)node[anchor=south] {ET Increasing};
        % \draw[-latex, line width=2pt] (0.83,0.725) --++(0.0in,-0.3in)node[anchor=north] {ET Decreasing};
        % \draw[dashed,-latex, line width=5pt] (0.8,0.75) --++(0.0in,+0.5in)node[anchor=north] {ET Decreasing};
    \end{scope}
    % \node[align=center,red,font={\large}] at (image.southwest) {ET decreasing};
  \end{tikzpicture}
\end{figure}
\begin{textblock*}{0.25\textwidth}(0.02\textwidth,0.3\textheight)
  \textcolor{magenta}{Dashed line} gives response for potential evapotranspiration (PET).\\
\medskip
  Plants are crucial for land response!
\end{textblock*}
\end{frame}

\section{Test Theory}
\begin{frame}{The theory seems nice, but we need to test with data!}
  \begin{itemize}
  \item Introduce a free uncertainty parameter \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize to Penman Monteith:
    \[ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6\; \textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}} \; \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right) }{ \Delta + \gamma}\]
  \item At each observation from FLUXNET (56 sites) calculate a unique \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize:
    \[\textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}}(t, \text{site}) = f(ET_{obs})\]

    \item Then propagate uncertainty forward by including \Large $\textcolor{magenta}{\boldsymbol{\sigma}}$ \normalsize in the derivative:
    \end{itemize}
      \[\frac{\partial \;  ET}{\partial \; VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} -  \frac{\gamma c_s }{1.6 \; R*\; \textcolor{magenta}{\boldsymbol{\mathlarger{\sigma}}} \; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]
\end{frame}

\begin{frame}{Test theory with FLUXNET data - the good}
  \begin{columns}
    \begin{column}{0.33\textwidth}
      \includegraphics[width=1.2\textwidth]{DBF_box.pdf}
    \end{column}
    \begin{column}{0.33\textwidth}
      \includegraphics[width=1.2\textwidth]{ENF_box.pdf}
    \end{column}
    \begin{column}{0.33\textwidth}
      \includegraphics[width=1.2\textwidth]{CSH_box.pdf}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Test theory with FLUXNET data - the bad}
    \begin{columns}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{GRA_box.pdf}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[width=\textwidth]{CRO_box.pdf}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}{Extra slide - is theory VPD$_{crit}$ optimum?}
  \includegraphics[width=0.7\textwidth]{for_agu.pdf}
\end{frame}
